---
title: "Assignments for Crim 250"
output:
  html_document:
    toc: yes
    toc_float: yes
    collapsed: no
    number_sections: no
    toc_depth: 1
  pdf_document:
    toc: no
    toc_depth: '1'
---


This page will contain all the assignments you submit for the class.

This page will contain all the assignments you submit for the class.



### Instructions for all assignments

I want you to submit your assignment as a PDF, so I can keep a record of what the code looked like that day. I also want you to include your answers on your personal GitHub website. This will be good practice for editing your website and it will help you produce something you can keep after the class is over.

1. Download the Assignment1.Rmd file from Canvas. You can use this as a template for writing your answers. It's the same as what you can see on my website in the Assignments tab. Once we're done with this I'll edit the text on the website to include the solutions.

2. On RStudio, open a new R script in RStudio (File > New File > R Script). This is where you can test out your R code. You'll write your R commands and draw plots here.

3. Once you have finalized your code, copy and paste your results into this template (Assignment 1.Rmd). For example, if you produced a plot as the solution to one of the problems, you can copy and paste the R code in R markdown by using the ` ``{r} ``` ` command. Answer the questions in full sentences and Save.

4. Produce a PDF file with your answers. To do this, knit to PDF (use Knit button at the top of RStudio), locate the PDF file in your docs folder (it's in the same folder as the Rproj), and submit that on on Canvas in Assignment 1.

5. Build Website, go to GitHub desktop, commit and push. Now your solutions should be on your website as well.






# Assignment 1

**Collaborators: none. **

This assignment is due on Canvas on Monday 9/20 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.


### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.

```{r}
# install.packages("datasets")
library(datasets)
```

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

```{r}
USArrests
dat <- USArrests
```
Answer: It is useful to rename the data set because it makes it easier to remember the name we give it to use in code commands later on in the assignment. It helps to separate the base R package dataset and turn it into our own dataset to use with the following work. 

### Problem 2

Use this command to make the state names into a new variable called State. 

```{r}
dat$state <- tolower(rownames(USArrests))
```

This dataset has the state names as row names, so we just want to make them into a new variable. We also make them all lower case, because that will help us draw a map later - the map function requires the states to be lower case.


List the variables contained in the dataset `USArrests`.

```{r}
list(dat)
head(dat)
names(dat)
summary(dat)

```

### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

Answer: Murder is a quantitative variable, because there are measured numerical values representing murder rates for each state. 

What R Type of variable is it?

Answer: Murder is a character variable, as seen in the summary (dat) description in the code above.


### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

Answer: This dataset contains the number of Murders, Assaults, and Rapes in each of the 50 states in the year 1973. It also shows us how many people lived in an urban area each state in that year, demonstrated by "UrbanPop."

### Problem 5

Draw a histogram of `Murder` with proper labels and title.

```{r}
# histogram of Murder
hist(dat$Murder, main="Histogram of Murder", xlab="Murder Rate", ylab="Frequency")

```

### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

```{r}
#Summary of Murder
summary(dat$Murder)
```
Answer: The mean of Murder is 7.788, the median of Murder is 7.250. Mean represents the average number of murders between the 50 states in 1973, whereas mean is the middle number of murders if the quantities were lined up in numerical order. 
Quartiles are 4 approximately evenly sized groups with the data ordered from least to greatest. R most likely gives us Q1 and Q3 because the values between these two represent the middle 50% of the data, and provide more context for the spread of the mean. 

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r}
#histogram of Assault
hist(dat$Assault, main="Histogram of Assault", xlab="Assault Rate", ylab="Frequency")
#Summary of Assault
summary(dat$Assault)

#Histogram of Rape
hist(dat$Rape, main="Histogram of Rape", xlab="Rape Rate", ylab="Frequency")
#Summary of Rape
summary(dat$Rape)

#Histogram of Murder, Assault, Rape
par(mfrow=c(3,1))
hist(dat$Murder, main="Histogram of Murder", xlab="Murder Rate", ylab="Frequency")
hist(dat$Assault, main="Histogram of Assault", xlab="Assault Rate", ylab="Frequency")
hist(dat$Rape, main="Histogram of Rape", xlab="Rape Rate", ylab="Frequency")
```

What does the command par do, in your own words (you can look this up by asking R `?par`)?

Answer: the command par is used to set certain parameters within the data given to it. In this case, we used the par function to tell R that we wanted to plot 3 character vectors on histograms together in one plot.

What can you learn from plotting the histograms together?

Answer: Plotting the 3 histograms together allows us to easily visually compare the frequency of murders, assaults, and rapes to see which crime was most common among the 50 states in the year 1973. 
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

Run this code:

```{r, fig.width = 7.5, fig.height = 4}
library('maps') 
library('ggplot2') 

ggplot(dat, aes(map_id=state, fill=Murder)) + 
  geom_map(map=map_data("state")) + 
  expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```

What does this code do? Explain what each line is doing.

Answer: These lines of code are generating a dataframe that shows the map of the US,  breaking up the information from the variable Murder into sections based on state. The lines also color-code the map based on the number of murders that occurred in that state. 

$$\\[2in]$$







# Assignment 2 

(Coming soon)


